---
title: "p8105_hw1_kiu2103"
output: github_document
date: "2025-09-18"
author: "Kaveri Uberoy"
---
Loading information and libraries for problems 1 and 2. 

```{r}
library(moderndive)
library(tidyverse)
data("early_january_weather")
```

The variables in the `early_january_weather` dataset are: `r paste(names(early_january_weather), collapse = ", ")`.

The size of the `early_january_weather` dataset is `r nrow(early_january_weather)` rows and `r ncol(early_january_weather)` columns.

The mean temperature for the early January weather dataset is `r mean(early_january_weather[["temp"]], na.rm = TRUE)`.

```{r}
library(ggplot2)

ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(title = "Temperature vs Time (Colored by Humidity)",
       x = "Time (in hours) from January 1st morning to January 14th night",
       y = "Temperature (°F)",
       color = "Humidity") +
  theme_minimal()

ggsave("scatterplot_temperature_vs_time.png", width = 8, height = 6)
```

Problem 2. Making a dataframe. 

```{r}
new_df = 
  tibble(
    random_samp = rnorm(10),
    logical_vector = rnorm(10) > 0,
    char_vector  = letters[1:10],
    factor_vector = factor(sample(c("low", "medium", "high"), 10, replace = TRUE), ordered=TRUE)
  )
```

Now, try to take the mean of each. 
```{r}
mean(pull(new_df, random_samp)) # Works, because it's numeric
mean(pull(new_df, logical_vector))# Works, logical values become either 0 or 1
mean(pull(new_df, char_vector)) # Will give an error, because characters don't automatically get converted to numeric
mean(pull(new_df, factor_vector)) # Will give an error, because factors don't automatically get converted to numeric
```

Try again after trying to convert to numeric 

```{r}
# Logical to Numeric (0 = FALSE, 1 = TRUE)
mean(as.numeric(pull(new_df, logical_vector)))
# Character with as.numeric() returns NA for each element, does not work
mean(as.numeric(pull(new_df, char_vector)))  
# Factor with as.numeric() returns the corresponding value for each factor
mean(as.numeric(pull(new_df, factor_vector)))
```